paper

### 2019-01-09 Report/paper.tex
pushing new graphs
push references

\begin{abstract}
    \#TODO write some sentence about general motivation for bibliometric
analysis We explore ways to extract information from ICLS and CSCL
papers to gain an understanding about the community formed by conference
participants and what research is being conducted within the community.
After describing how to extract data from the original dataset in
suitable format for analysis we perform an exploratory data analysis
after which we derive insights into the community using social network
analysis and {[}Nour: NLP etc{]}. \#TODO write a few centences about the
most central findings
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum.

@@

    As a starting point we have papers of 4 years spanning 2015 to 2018 of
the proceedings of ICLS, which is held in even years, respectively CSCL,
held in odd years. The overall number of papers per year is shown in
figure \(\ref{fig1}\). The provided raw dataset consists of papers in
pdf format as well as associated metadata in xml format. To extract
information from the dataset we first need to convert the data a format
that is more adapt to the task of data exploration and handling with a
variety of external libraries. We store the data we extract from the
provided dataset in csv format, as this makes it accessible to any
scientist and is unlikely to become an obsolete any time soon.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_8_0.png}\end{center}
        \caption{Distribution of papers published over the years}
        \label{fig1}
    \end{figure}

@@

We extract the xml metadata into a csv file containing the information
associated to different tags in its columns using the \emph{lxmltree}
python library. The provided xml does not exactly follow the xml
standard, hence we handle errors by iteratively parsing the xml tree. We
then add columns that reference the name of the source file and the
order in which an author name occurs in the xml file. From the
associated citation string we add in a shortened version of author
names, and the ``author names (year) title'' string that we later use to
identify a paper. Having extracted the metadata we then clean it by
unifying names as slight spelling differences exist within the same
author name. This is done by splitting up names based on whitespace and
commas. If the intersection of the set of substrings of two distinct
names contains at least 2 names of length at least 3 we consider these
two names to be identical. Using this we reduce the number of authors in
the dataset from 1951 down to 1879. To unify the names further, we need
to make sure to not overmerge by labeling to different authors as having
the same name. We do this by constructing the graph of co-authorship and
then only considering names of authors that are in the same neighborhood
of a node on the graph. Then we use the difflib library to check for
similarity of two names and merge them if they reach a the threshold of
0.8. This method again reduces the number of authors in the dataset from
1879 to 1951.

@@

To extract information from the papers themselves we need to first
extract the text from the pdfs. Various parsers performing this task
exist. However, many parsers {[}pdfminer, PyPDF2{]} claiming to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is poppler https://poppler.freedesktop.org/. We
note that due to inconsistencies with numbering, some of the pdfs were
not extracted correctly from the proceedings which leads to text files
that are either empty or do not contain the text of the paper referenced
in the metadata.

The advantage of papers is that they are structured fairly regularly.
From ISLS Guildines \ref{ISLS_Author_Guidelines_Oct2014.pdf} we know
papers the papers in our dataset should be structured. We use this to
our advantage in the following sections when extracting information from
the paper text.

@@

We know that each valid paper should contain a reference section at the
end listing all the references cited in the paper. While there are tools
being developed to extract the reference section directly from pdfs,
notably {[}scholarcy{]}, we found that none of them produced
satisfactory results as of the time writing this report. Hence we resort
the using regular expressions and our knowledge about how references
should be presented within the paper.

To get the reference section we use a regular expression to find the
`Reference' header and cut of at the end if an Acknowledgement or
Appendix section is present. Python has two modules that implement
regular expressions \emph{re} and \emph{regex}. We used regex as it
allows to better handle unicode strings by having unicode expressions.
This is important as many authors have names that do contain non-ascii
characters. To get the individual reference we tested many ways to split
up the lines, but ultimately, the most robust way was to split up the
section at each new line and the glue the references back together if
the lines satisfy a set of conditions.\\
- The underlying idea is that APA references, the standard referencing
format, all have the same initial structure: Authors (year) title, other
identifying information. We thus check that the beginning of each
reference contains ``Authors (year)''. It is relatively easy to check
for this using regular expressions as Author names and year have a given
structure. If a string does not contain such a section, then it must be
part of an other reference, which is lies before in the list of
reference substrings.

Consider that the references strings are structured the following way
{[}\ldots{}., reference\_part\_i, reference\_part\_i+1, \ldots{}.{]}
Algorithm:

\[
1. Merge names that were cut up. We detect them by checking against the regular expression: __ 
2. Check ending for invalid sentence endings:  reference_part_i ends in comma, :, whitespace uppercase followed by point. If condition holds reference_part_i+1
3. Check beginning for invalid reference starts: not starting in a letter. Merge with reference_part_i if condition holds.
4. Check if a  invalid reference_part_i+1 consists only of the authors names. If so Merge with reference_part_i
5. Move up all sentences that do not contain author names / start of a citation at the beginning 
6. Move up all sentences that do not have a valid citation start
\] After this we validate the returned split by looking at the overall
length and the initials, as well as checking that a random sample
returns satisfactory splitting accuracy. We find that the above method
of extracting the references produces good results for proper APA
citations and on APA citations outperforms Scolarcy.

@@

Now we consider the information present in the section before the
abstract, which contains the authors names, their email and the
institution they are affiliated with. Affiliation with institution can
give us a lot of insight where people are working and gives use a way
estimate where they are from. To go from this section to an exact
affiliation with an institution a multifaceted approach was used.

@@

The most robust way to get a persons institution affiliation is to use
their email. Emails can easily be extracted from the header section
using a regular expression and given a domain-university mapping {[}{]}
give us an onto mapping from domain to institution. The additional
advantage of this method is that using the additional metadata
associated to each university in the mapping, we also access information
about the country of the university.

@@

The header section itself contains information on the institution, but
as different was to list authors are accepted, notably two ways {[}fig
\_ a and b{]}. Hence, we first detect the structure of the header
section before extracting the association. We tag the substring present
in this section using a combination of named entity detection, using the
implementation provided by spacy, simple matching of the names of
authors and institutions we extracted from the metadata respectively
from Approach 1 and tagging sections containing ``emails'' using regex ,
to detect where in the header the email is. Then, taking the tagged
section, we detect which subsections correspond to which
author-affiliation listing type, and extract the author email -
institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping and many
participants use gmail and similar providers as their official
correspondence email we are unable to acquire this information for many
participants. But, Approach 2 may contain differences in spelling in the
institution name, which need to be merged. Thus we combine both
approaches, and use the corresponding institutions in Approach 1 to
unify the institutions obtained in Approach 2. As institution
affiliation does not allow us to automatically associate a country to
the institution, we need to get this through additional means. A first
step is to use the top level domains present in the emails to associate
a country of origin to the participants \& their associated
institutions. Alongside this, to plot the community we augment the
dataset with geolocation information, having this information we can
then easily associate a country to the institution.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- it was found that crawling
the english, french and german version of Wikipedia (not using the API,
or the API wrapper, as this does not give access to the desired
information) and using \emph{BeautifulSoup} the get the coordinates
present in the article was a quick and accurate way to get the location
as the website directs to the relevant article on the institution even
when misspellings are present. But not all institutions, even well known
ones such as Tokyo University, have their coordinates listed on their
Wikipedia. To get the remaining locations we use the geopy, which is a
client for geocoding web services. These results however are more error
prone and will always return a result. Then, to get the country we can
then again use geopy. The resulting data containing paper identifier,
the authors email, it's associated domain, order in which the author was
cited, the institution, country, Longitude and Latitude are saved to
csv. We use the paper identifier and author order to associate this data
with the metadata to link information about authors with the papers they
authored.

@@

\begin{itemize}
\item
  We used \textbf{\emph{beautiful soup}} and \textbf{\emph{string}}
  libraries to only keep the text that we needed without any special
  character.
\item
  we applied \textbf{\emph{regular expression}} to remove numbers
  because we thought that numbers were not that important.
\item
  We put all the words to lower cases
\item
  We removed the words that were connecting parts of a sentence rather
  than showing subjects, objects or intent. For us these words were not
  that important and were only going to increase the bag of words length
\item
  In some methods we used \textbf{\emph{Lemmatization}} and
  \textbf{\emph{stemming}} and in other methods we didn't because we
  wanted to do a mapping with \textbf{\emph{GloVe}} vocabulary which we
  will explain later on.
\end{itemize}

@@

    \textbf{\emph{Lemmatization}} and different form of
\textbf{\emph{Stemming}} were only used with \textbf{\emph{TF-IDF}}
while doing the Keywords extraction, otherwise we prefered keeping the
words as they are because we were doing a mapping with a pretrained
stanford word embedding and we did not wand to loose interpretability.

    For the Keywords extraction we used three methods. The first one whe is
a trivial one is to search in the pdf itself for keywords and extract
them, this method is the most accurate but we found out that for 492
papers we had no keywords included. So we deciced to use two methods for
keywords extraction. The first one is \textbf{\emph{TF-IDF}} and the
second one is \textbf{\emph{Rake}}.

@@

    Once we have all the keywords from the three methods, we tried to select
the most accurate keywords. So we created a method that selects the
intersection of TF-IDF keywords and the Rake keywords with the keywords
of the authors (if there are any keywords in the text), if the
intersection is empty and we have the keywords in the text we only take
the keywords of the text. If we don't have any keywords we take the
intersection of the TF-IDF and the Rake. If again, the intersection is
empty then we only take the TF-IDF keywords because they were closer to
the text keywords than the Rake Keywords.

    Once we have our final keywords, we created a method that goes into the
metadata and put assign to every document a publication year.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_25_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we selected distinct words for each year, it means that we
selected for each year the words that were present for that year and
never for the others. We get the graph below. As we can see from the
graph, the most used word in the papers, more than 10\% of 2015 papers
uses the word `peer', more than 6\% of the papers uses the words
`building', `embodied' in 2016. In 2017 the most frequent keyword on is
`reponsive' and in 2018 `tools' is present with more than 6\%.

@@

    \hypertarget{co-authorship-graphs}{%
\subsection{Co-authorship graphs}\label{co-authorship-graphs}}

\hypertarget{interactive-map-visualizations}{%
\subsection{Interactive Map
visualizations}\label{interactive-map-visualizations}}

>>>>

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conferences
participants and of the research being conducted within the community.
After describing how to extract data from the original dataset we we
derive insights into the community using social network analysis and
{[}Nour: NLP etc{]}. We discover patterns of collaboration between
participants, their migration and the global spread of the community.
{[}Something about the insights in keywords/NLP/Co-Citation{]}
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The starting point of the project are the proceedings of the
International Conference of the Learning Sciences (ICLS), respectively
the International Conference on Computer-Supported Collaborative
Learning (CSCL). The two conferences are organized by the International
Society of the Learning Sciences (ISLS) and held in biennial alternation
with each other. At the time of writing the dataset contains 4 years of
the proceedings, staring from 2015 to 2018. While ICLS covers the entire
field of the learning sciences, CSCL focuses on learning through
collaboration with the help of communication technologies \cite{ICLS}.
To understand the community formed by the participants of the
conferences as well as to understand the research being conducted within
the community, we explore methods to bring insight into the following
aspects:

\begin{itemize}
\tightlist
\item
  the institutions and countries dominant in the conferences
\item
  the differences between ICLS and CSCL with respect to citations,
  contributors and publications
\item
  first insight into changes of the years with respect to citations,
  contributors and publications
\item
  collaboration network patterns with respect to countries, institutions
  and authors
\item
  the migration of participants across institutions
\item
  \ldots{} {[}Nour{]}
\end{itemize}

To this means we first present an extensive pipeline to process the
textual data and extract features relevant to bibliographic analysis
from the textual data. After this we perform a first bibliographic
analysis based on the extracted data.

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first convert the data to a format that is more
adapt to the task of data analysis and more compatible with a variety of
external libraries. We store all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_7_0.png}\end{center}
        \caption{Distribution of papers published over the years}
        \label{fig1}
    \end{figure}

@@

We extract the metadata from the xml files using the \emph{lxmltree}
python library. We associate the contents of each xml tag to a column in
csv file. As the xml files do not exactly follow the xml standard, we
handle errors by iteratively parsing the xml tree. For some files, the
section containing the keywords is malformed which can not be handled by
the parser. This is remedied by using regular expressions to match the
section contains the keywords. To each file we associate the name of the
source file and the order in which an author name occurs in the xml
file. Each author is associated one line in the XML file. From the
associated citation string we add to the first part of the citation
containing the author names, year and title. which we use to identify
references in other papers to that paper. Then we add a shortened
version of author names (i.e.~West, R.) by considering the author order.
Note that the short name extracted from the citation and the
corresponding long name (Robert West) in the metadata do at times not
match up. Thus the author order listed in the citation string does not
always correspond to the author order listed in the metadata. After
extraction, we clean the data by unifying names, as slight spelling
differences exist within the same author name (i.e.~Robert West, Robert
A. West). When unifying the names we need to make sure to not overmerge
by labeling to different authors as having the same name. First, we use
a strict condition to catch small differences in names, such as an extra
middle name. This is done by splitting up names into substrings based on
whitespace and commas. For two names in the dataset we check whether the
intersection of their sets of substrings contains at least 2 names of
length at least 3. If the condition is met we consider these two names
to be identical, and add the mapping from one name to the other to a
dictionary. Once all tuples of names (s.t. the first name appears before
the second one in order) are checked and added to the dictionary we
unify the names. Using this method we reduce the number of authors in
the dataset from 1951 down to 1879. Further, to merge names with
misspellings or people that use nicknames (Christian vs Chris) and avoid
over-merging we consider only names of authors that have similar
collaboration patterns. To do this we build the graph of co-authors and
check names of authors that are in the same neighborhood of a node on
the graph (see methodology for further details). Then we use the
\emph{difflib} library to check for similarity of two strings and
consider a name to belong to the same person if they reach the threshold
of 0.8. Again names considered to belong to the same person are added to
the mapping dictionary which is then applied to all names. This method
again reduces the number of authors in the dataset from 1879 to 1951.

@@

To extract information from the papers themselves we need to first
extract the text from the pdfs. Various parsers performing this task
exist. However, parsers such as pdfminer, PyPDF2 that claim to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is \emph{poppler} \cite{poppler} which can be
installed and then used via the command line to extract the text
contained in the pdfs to \emph{txt} format. All methods describing how
to extract data from the papers work on the extracted \emph{txt} files.
The characters contained in the \emph{txt} files are not all encoded
uniformly, which means that characters that look the same to humans may
be encoded in different way . Hence, for many applications we need to
unify the formating. This is done using the \emph{unicodedata} module
and setting converting all characters to `Normal Form Composed' form .

A convenient property of papers is that they are structured fairly
regularly. The ISLS Author Guidelines \cite{guidelines} give us guidance
on how the papers in our dataset should be structured. We use this to
our advantage when extracting information from the paper text. While the
vast majority of papers in the dataset follow these guidelines,
exceptions do exist. We thus make sure to handle deviations from the
structure.

@@

We know that each valid paper contains a reference section at the end
listing all the references cited in the paper. While tools are being
developed to extract the reference section directly from pdfs, notably
Scholarcy \cite{scholarcy}, they did not produce satisfactory results.
That is they had issues splitting the references correctly and
categorizing parts of the reference accurately. This may occur because
they are trained to extract any reference format, and thus can not make
many assumptions on the structure. But in our dataset papers following
the guidelines should list their references in APA format. Hence using
regular expressions and background on how APA references gives us better
results than out of the box methods.

\textbf{Regular expressions in python:} Python has two modules that
implement regular expressions \emph{re} and \emph{regex}. We used regex
as --unlike the re module-- this module allows to handle strings with
non-ascii characters by having unicode expressions. As many authors have
names with non-ascii characters this is an important capability.

\textbf{To get the reference section for of a paper} we use a regular
expressions to find the reference header and determine the sections end
by checking for acknowledgement or appendix sections in the text file
previously extracted using poppler. Then, to extract individual
references we split the section into a list of substrings at each new
line character and glue substrings back together if they satisfy a set
of conditions. We found this method to be robust as it makes sure that
all properly formatted references are separated.

\textbf{To glue back substrings} we exploit the fact that that APA
references all have the same initial structure: \emph{Lastname F.M.
(when published) Title which may contain colons and other non-alphabetic
chars. Where published. Who published} \cite{apaformat}. We thus check
that the beginning of each reference contains ``Lastname F.M. (when
published)''. It is relatively easy to check for this using regular
expressions as author names and year have a given structure. If a string
does not contain such a section, then it must be part of an other
reference, which is lies before in the list of reference substrings.
Note that this beginning is also what is added to the metadata as
identifying string. We found that in practical application it is
important to first merge lines if they were clearly cut off, as papers
with many authors may have the beginning on multiple lines. We can find
split up lines by checking for lines starting with non-alphabetic
characters or starting in a sentence. We merge these with the line
above. We can identify lines by their ending as being split up, such as
if they end in a comma, colon or only contain author names. We merge
these lines with the lines below them. Only then can we check whether a
line contains a valid citation start. If it does not we merge it with
the line above it. After this merging process we validate the returned
reference split by looking at the overall length and the initials, as
well as checking that a random sample returns satisfactory splitting
accuracy. We find that the above method of extracting the references
produces good results for proper APA citations and on APA citations
indeed outperforms Scolarcy when it comes to splitting up the lines.
\emph{To then get the subcomponents of a reference} is rather trivial,
as we can again use the structure of APA citations. The string before
the first parenthesis contains the authors names, inside the parentheses
we find the year or some other reference to publication time, after
which the title follows until the first dot. Then a reference to the
publishing venue follows.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_10_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figA}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_10_1.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figA}
    \end{figure}

@@

Now we consider the section before the abstract, which contains the
authors names, their email and the institution they are affiliated with.
Affiliation with institution can give us a lot of insight where people
are working and gives use a way estimate where they are from. To go from
this section to an exact affiliation with an institution a multifaceted
approach was used.

@@

The most robust way to get a persons institution affiliation is to use
their email. Emails can easily be extracted from the header section
using a regular expression and given a domain-university mapping
\cite{mapping} give us an onto mapping from domain to institution. The
additional advantage of this method is that using the additional
metadata associated to each university in the mapping, we also access
information about the country of the university.

@@

The header section itself contains information on the institution, but
as different was to list authors are accepted, notably two ways as can
be seen in \ref{figA} \ref{figB}. Hence, we first detect the structure
of the header section before extracting the association. We tag the
substring present in this section using a combination of \emph{named
entity detection}, using the implementation provided by \emph{spacy},
simple string comparison matching of the names of authors and
institutions we extracted from the metadata respectively from Approach 1
and tagging sections containing ``emails'' using regex , to detect where
in the header the email is. Then, taking the tagged section, we detect
which subsections correspond to which author-affiliation listing type,
and extract the author email - institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping --no
companies will be in the mapping-- and many participants use gmail and
similar providers as their official correspondence email. However,
Approach 2 may contain differences in spelling in the institution name,
which need to be merged. Thus we combine both approaches, and use the
corresponding institutions in Approach 1 to unify the institutions
obtained in Approach 2. Furthermore, we later merge names by considering
the location of the institutions.

But first as with Approach 2 institution affiliation does not allow us
to automatically associate a country to the institution, we need to get
this through other means. A first step is to use the top level domains
present in the emails to associate a country of origin to the
participants and their associated institutions. Then, to plot the
community we augment the dataset with geolocation information. As a
further benefit, having this information we can then easily associate a
country to the institution for the emails that do not have a top level
domain indicating location and merge the names of institutions where
necessary based on location.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present. Requests were
sent using \emph{urllib3} and \emph{beautifulsoup} was used to parse
them get the coordinates present in the article. Regretfully not all
institutions, even well known ones such as Tokyo University, have their
coordinates listed on their Wikipedia. To get the remaining locations we
use the geopy, which is a client for geocoding web services. These
results however are more error prone and will always return a result. It
is thus recommended to verify the correctness of the results. To get the
country we can then again use geopy. The resulting data containing paper
identifier, the authors email, its associated domain, order in which the
author was cited, the institution, country, Longitude and Latitude are
saved to a csv file. We use the paper identifier and author order to
associate this data with the metadata to link information about authors
with the papers they authored.

@@

\begin{itemize}
\item
  We use \textbf{\emph{beautiful soup}} and \textbf{\emph{string}}
  libraries to only keep the text that we needed without any special
  character.
\item
  we apply \textbf{\emph{regular expression}} to remove numbers because
  we thought that numbers were not that important.
\item
  We put all the words to lower cases
\item
  We remove the words that were connecting parts of a sentence rather
  than showing subjects, objects or intent. For us these words were not
  that important and were only going to increase the bag of words length
\item
  In some methods we use \textbf{\emph{Lemmatization}} and
  \textbf{\emph{stemming}} and in other methods we don't because we
  wanted to do a mapping with \textbf{\emph{GloVe}} vocabulary which we
  will explain later on.
\end{itemize}

@@

    \textbf{\emph{Lemmatization}} and different form of
\textbf{\emph{Stemming}} were only used with \textbf{\emph{TF-IDF}}
while doing the Keywords extraction, otherwise we prefer keeping the
words as they are because we are doing a mapping with a pretrained
stanford word embedding and we do not want to loose interpretability.

    For the Keywords extraction we use three methods :

The first one whe is a trivial one is to search in the pdf itself for
keywords and extract them, this method is the most accurate but we found
out that for 492 papers we had no keywords included. So we deciced to
use two methods for keywords extraction. The first one is
\textbf{\emph{TF-IDF}} and the second one is \textbf{\emph{Rake}}.

@@

    Once we have all the keywords from the three methods, we try to select
the most accurate keywords. So we create a method that selects the
intersection of TF-IDF keywords and the Rake keywords with the keywords
of the authors (if there are any keywords in the text), if the
intersection is empty and we have the keywords in the text we only take
the keywords of the text. If we don't have any keywords we take the
intersection of the TF-IDF and the Rake. If again, the intersection is
empty then we only take the TF-IDF keywords because they were closer to
the text keywords than the Rake Keywords.

    Once we have our final keywords, we create a method that goes into the
metadata and assign to every document a publication year.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_26_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we select distinct words for each year, it means that we select
for each year the words that were present for that year and never for
the others. We get the graph below. As we can see from the graph, the
most used word in the papers, more than 10\% of 2015 papers uses the
word `peer', more than 6\% of the papers uses the words `building',
`embodied' in 2016. In 2017 the most frequent keyword on is `reponsive'
and in 2018 `tools' is present with more than 6\%.

@@

To cluster the documents we decided to use two different approaches. The
first one is Natural Langage Processing (NLP) and the second one is a
network analysis. The most accurate one is the network clustering using
the references graph.

\hypertarget{approach-1-natural-langage-processing-nlp}{%
\subsubsection{Approach 1 : Natural Langage Processing
(NLP)}\label{approach-1-natural-langage-processing-nlp}}

\hypertarget{baseline-tf-idf-combined-with-k-means}{%
\paragraph{Baseline : TF-IDF combined with
K-means}\label{baseline-tf-idf-combined-with-k-means}}

With the natural langage processing, we start with a naive method which
consists of clustering the documents using only TF-IDF. Each document
has a vector representation with values for each word present in the
document depending on the frequency of the word in this document
regarding its frequency in all the other documents. So if a word is a
keyword to the document and is not a common word than its value will be
more important than the other words in the vector of the document. Once
we have our vector representation, we select the max features parameter
as 2500, it means that we will only consider the top max\_features
ordered by term frequency across the corpus. On this word representation
we will run K-means.

\textbf{\emph{K-means}} clustering is a method used for clustering
analysis, especially in data mining and statistics. It aims to partition
a set of observations into a number of clusters (k), resulting in the
partitioning of the data into Voronoi cells. It can be considered a
method of finding out which group a certain object really belongs to.

This method does not give good results as we can see it in the graph
below all the vectors are together there is no clear separation between
the documents. So we decide to use other methods using word embeddings
and properties of TF-IDF.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_30_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    \hypertarget{second-method-average-of-glove-vectors-combined-with-k-means}{%
\paragraph{Second method : Average of GloVe vectors combined with
K-means}\label{second-method-average-of-glove-vectors-combined-with-k-means}}

The second method was to use pre-trained words embeddings. We decided to
use \textbf{\emph{GloVe}} library pre-trained on Wikipedia dataset.

\textbf{\emph{GloVe}}, coined from Global Vectors, is a model for
distributed word representation. The model is an unsupervised learning
algorithm for obtaining vector representations for words. Training is
performed on aggregated global word-word co-occurrence statistics from a
corpus (Wikipedia), and the resulting representations showcase
interesting linear substructures of the word vector space. It is
developed as an open-source project at Stanford.{[}1{]}

We construct a dictionnary from the words embeddings mapping each word
of a document to its vector representation, than we average all the
vectors of each document. Since each words has its representation in
space, the average of all the words for each document will lead to a
vector representing a document.

Once we get a document vector representation, we apply to it K-means.
The results are better, but we have only two clusters. If we increase
the number of clusters than there is no clear separation anymore
separation.

    \hypertarget{third-method-weighted-average-of-tf-idf-glove-combined-with-k-means}{%
\paragraph{Third method : Weighted average of TF-IDF + GloVe combined
with
K-means}\label{third-method-weighted-average-of-tf-idf-glove-combined-with-k-means}}

With this method, we did not use the parameter max features of the
TF-IDF vector, we took all the vocabulary generated by all the documents
and we created a method that sorts the document frequency vector, take
the 500 first words and do a weighted average of the words embedding
vectors with the TF-IDF score.

Once we get the weighted average vector, and since we are using a words
embedding of 50 dimentions. We decide to use feature reduction technique
to reduce the number of dimensions and represent our vectors in an other
vector space where all the features are orthogonal to each others.

To apply the feature reduction to our vectors we used
\textbf{\emph{Principal Component Analysis (PCA)}} algorithm.

\textbf{PCA} is a statistical procedure that uses an orthogonal
transformation to convert a set of observations of possibly correlated
variables (entities each of which takes on various numerical values)
into a set of values of linearly uncorrelated variables called principal
components. If there are \textbf{\emph{n}} observations with ****p****
variables, then the number of distinct principal components is
\textbf{\emph{min(n-1,p)}}. This transformation is defined in such a way
that the first principal component has the largest possible variance
(that is, accounts for as much of the variability in the data as
possible), and each succeeding component in turn has the highest
variance possible under the constraint that it is orthogonal to the
preceding components. The resulting vectors (each being a linear
combination of the variables and containing n observations) are an
uncorrelated orthogonal basis set. PCA is sensitive to the relative
scaling of the original variables.

Once we get our new vector representation, we apply K-means and we get
the best separation, with a clear separation of 4 clusters as we can see
it in the figure below.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_34_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    \hypertarget{approach-2-references-network-analysis}{%
\subsubsection{Approach 2 : References Network
Analysis}\label{approach-2-references-network-analysis}}

For the references graph we choose to construct 4 different graphs to be
able to extract the more information we can. Our first purpose is to be
able to cluster the documents that are about the same subject.

    \hypertarget{first-reference-graph-intersection-between-referenced-documents}{%
\paragraph{First reference graph : Intersection between referenced
documents}\label{first-reference-graph-intersection-between-referenced-documents}}

Our first graph is constructed using the references of each document. It
is constructed as follow, for each two documents, if there is an
intersection between the references of both documents then there is a
link. The weights of the edges are the length of the intersection set.
The intuition behind is that the more the weight is high, the more the
documents are likely to be about the same subjects since they are citing
the same documents. From this graph, we create a subgraph of it, we
filter the edges keeping only those who have a weight strictly higher
than 3. It means that, each node of the subgraph, is related to at least
one other node with 4 same documents referenced.

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}9}]:} <IPython.core.display.HTML object>
\end{Verbatim}

    The resulted graph, is an undirected graph with 805 nodes and 6699
edges. On this graph we extract two main information : * Strongly
connected components to be able to detect clusters and related
documents. * In\_degree in order to check if there are documents that
are related to a lot of documents and those documents would be more
general than others, and won't necessarly belong to one specific
cluster.

    \hypertarget{second-graph-relations-between-references}{%
\paragraph{Second graph: Relations between
references}\label{second-graph-relations-between-references}}

The second graph is more about the referenced documents and not about
the documents themselves. The aim of this graph is to see if there are
documents that are more likely to cited together. If it is the case,
then these documents are likely to be about the same subject. It is a
weighted graph constructed as follow: for each document, we create a
link between the references of the document. We create a dictionnary of
links, if the link is already in the dictionnary then we increase the
weight of the link. So for instance, if two references are present
together in the 5 distinct documents, then the weight of the edge would
be of 5. We obtain a graph of 10845 nodes and 148594 edges.

    \begin{Verbatim}[commandchars=\\\{\}]
Uploading 7403 kB. This may take a while{\ldots}

    \end{Verbatim}

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}11}]:} <IPython.core.display.HTML object>
\end{Verbatim}

    We can directly select from the interactive graph the nodes with degree
higher than 100, these nodes represent the more cited documents, but we
will construct a subgraph to specificly tackle these important
referenced documents. On this graph, to be able to detect similar
documents we also run Strongly connected components algorithm. On top of
it, we inspected the documents with higher weights, we will explain more
about what we extracted from this graph in the result section.

    \hypertarget{third-graph-title-to-find}{%
\paragraph{Third graph: ??? Title to
find}\label{third-graph-title-to-find}}

With this graph we aim to find what are the most cited documents, what
are the documents that are the most influencial and how citations evolve
over time.

This is a directed graph constructed in this way. For each document, if
the document cites another document than there is a link between the
document and the cited one. We get a graph of 11650 nodes and 13243
egdes. We also construct a subgraph of this graph, which is a
co-citations in conferences, but because a lack of data, we find only
three documents that cites other documents.

On this graph, we run \textbf{\emph{PageRank}} algorithm to be able to
extract the most influential documents.

\textbf{PageRank (PR)}* is an algorithm used by Google Search to rank
web pages in their search engine results. In our case, It measure
PageRank is a way of measuring the importance of referenced papers.
Compared to the in-degree analysis we get almost the same results since
we have no out degree for almost every node.

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}13}]:} <IPython.core.display.HTML object>
\end{Verbatim}

    \hypertarget{co-authorship}{%
\subsection{Co-authorship}\label{co-authorship}}

To get insight into collaborations within the community we build graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. First, we
built a co-authorship graph based on individual authors. Each node
represents an author, and two nodes are connected if the authors
collaborated on a paper. Other works \cite{cheong2009social},
\cite{hesford2006management} have used a directed graph that only adds a
link from the first author of the paper to all other authors. However,
as we are more interested in collaboration patterns and less in finding
the most influential first author, we build an undirected graph with
edges between all co-authors. On the downside, yields a very dense graph
that can not directly be interpreted. Hence, we prune the initial graph
by removing nodes that do not satisfy additional conditions. Intuitively
it is clear that a singular collaboration on a paper does not
necessarily indicate that two people even know each other. A first
condition was to only link two nodes together if the authors
corresponding to each node have appear as co-authors in more than one
conference. The graph produced by this approach is much more sparse and
contains much fewer authors. Additionally all connected components
contain fewer than 50 authors, making it possible to consider each
author individually. Such a component is shown in figure \ref{fig_plot}.
A second alternative approach was to consider the strength of bounds. In
the graph authors are only linked if they have written at least 2 papers
together. This method yields a much larger component --containing 441
authors-- as this second condition is weaker. To get an understanding of
author to author relations in this component we need to further prune
away parts of the graph that are not insightful. When we look at
individual author within the community, we are interested well
collaborative individuals. Hence we remove nodes that have collaborated
with fewer than 3 connections to other authors. The resulting graph can
be seen in figure \ref{author_collab_large}. Details creating such a
plot can be found in the section dedicated to this below.
